---
title: "Wyndsor Glass Example - Post Optimality Analysis"
output:
  html_document:
    df_print: paged
  html_notebook:
    highlight: textmate
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```

***

This notebook contains the code for the examples in Chapter 5. Specifically, for postoptimality analysis.

```{r}
library(lpSolveAPI)
```

***

Read the WGC example from file

```{r}
x <- read.lp("Wyndsor.lp") # create an lp object x
x                          # display x
```

Solve the lp model
```{r}
solve(x)
get.objective(x)        # get objective value
get.variables(x)        # get values of decision variables
get.constraints(x)      # get constraint RHS values
get.sensitivity.rhs(x)  # get shadow prices
get.sensitivity.obj(x)  # get reduced cost
```

Note that the shadow prices are expressed here under $duals. For example, the shadow prices for  constraint 1 is 0, for 2 is 1.5, and for 3 is 1. The last two values refer to the non-negativity constraints. We also have valid ranges for shadow price calculations. Those are given under $dualsfrom and $dualstill. Note for constraint 2 it is from 6 to 16, exactly what we found through the graphical method

The reduced costs are expressed here until $objfrom and $objtill. The output here indicates that as long as the contribution to profit for product 1 is between 0 and 7.5, it can remain in the solution. Similarly, 2 <= C2 <= infinity, where c2 is the contribution to profit for x2, x2 will be part of the final solution. If c2 < 2, then x2 will no longer be part of the final solution.